---
title: 'hw04: Eigenvalues'
author: 'Kenji Sato^[Kobe University. Email: mail@kenjisato.jp]'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    includes:
      in_header: asset/latex/custom.sty
    keep_tex: no
    latex_engine: pdflatex
    number_sections: yes
    template: null
    toc: no
  bookdown::html_document2: default
documentclass: scrartcl
---

# Overview

## Purpose {-}

- Study lists in R
- Compute eigenvalues and eigenvectors in R.

## Instructions {-}

In this assignment, you will

- clone the assignment repository and make a working branch (eg. `solution` branch);
- solve the problems in Section \@ref(problems); 
- write the solutions in `solution.Rmd`; 
- commit `solution.Rmd` and `solution.pdf`; and 
- open a Pull Request.
 

# Lists in R

## List basics 

A list in R is a vector that can contain another vector. It is also called as a recursive vector. 
Unlike atomic vectors, a list can have any type of objects. `list()` function creates a list.

```{r}
(x <- list(3.1, -2.2, 0.3))
```

Subscription `[` works differently for lists. 

```{r}
x[1]
```

`x[1]` is again a list. To get `3.1` you need to use `[[`.

```{r}
x[[1]]
```

Here is a more practical example.

```{r}
y <- list(1.1, c(2.1, 2.2), c(3.1, 3.2, 3.2))
y
```

Now you must be able to get `3.2`. 

```{r}
y[[3]][1]
```

How about this? Each element of `z` is a list.

```{r}
z <- list()
for (i in seq_along(y)) {
  z[i] <- list(y[[i]])
}
z
```

To get `3.2`, do this.

```{r}
z[[3]][[2]]
```

## Named elements

Elements in list can have names.

```{r}
w <- list(a = 1, b = 2)
w
```

There are several ways to get elements from `w`. To get sublists:

```{r}
w[1]
```
```{r}
w["a"]
```
To get an element:
```{r}
w[[1]]
```

```{r}
w$a
```
or 
```{r}
w[["a"]]
```

## List as information store

Since list can contain any vector as elements, you can use lists as an information store.

```{r}
alice <- list(
  name = "Alice",
  age = 18,
  grades = c(100, 100, 95)
)

bob <- list(
  name = "Bob",
  age = 19,
  grades = c(90, 100, 100)
)
```

```{r}
students <- list(alice, bob)
students
```

To get Alice's grade for the second course, do something like this.

```{r}
students[[1]]$grades[2]
```

# Eigenvalues and eigenvectors

Now we are ready to compute eigenvalues.

```{r}
set.seed(10903)
A <- matrix(rnorm(9), nrow = 3)
A
```

```{r}
eig_A <- eigen(A)
eig_A
```

`eigen(A)` returns a list of eigenvalues and eigenvectors. 
`eigen(A)$values` is an atomic vector that consists of eigenvalues of `A`. `eigen(A)$vectors` is a matrix of eigenvectors. The $n$-th column of `eigen(A)$vectors` is an eigenvecotr that corresponds to the $n$-th element of `eigen(A)$values`.

```{r}
lambda = eig_A$values
V = eig_A$vectors

all.equal(A %*% V[, 2], matrix(lambda[2] * V[, 2]))
```

`V[, 2]` is the atomic vector consisting of the second columns of `V`. 

The above computation corresponds to the following identity.  
$$
Av_2 = \lambda_2 v_2
$$
with the matrix of eigenvectors
$$
V = \begin{bmatrix}
  v_1 & v_2 & v_3
\end{bmatrix}
$$
The following identidy is true.
$$
A \begin{bmatrix}
  v_1 & v_2 & v_3
\end{bmatrix}
=
\begin{bmatrix}
  v_1 & v_2 & v_3
\end{bmatrix}
\begin{bmatrix}
  \lambda_1 & & \\
  & \lambda_2 & \\
  & & \lambda_3
\end{bmatrix}
$$
In R, this can be written as
```{r}
all.equal(A %*% V, V %*% diag(lambda))
```

# Exercise

Let $A$ be a square matrix, $\Lambda$ a diagonal matrix consisting of $A$'s eigenvalues 
and $V$ the matrix of eigenvalues for which $AV = V\Lambda$ holds.
If $A$ is diagonalizable, 
$$
A = V\Lambda V^{-1}
$$
holds.

Let `A` be a $1000 \times 1000$ matrix. 

```{r}
set.seed(1093)
n <- 100
A <- matrix(rnorm(n * n), nrow = n)
```

Verify that $A = V\Lambda V^{-1}$.


## Solution {-}

Here is a sample code.

```{r}
r <- eigen(A)
V <- r$vectors
Lambda <- diag(r$values)

all.equal(A + 0i, V %*% Lambda %*% solve(V))
```


\pagebreak

# Problems

Let 
$$
B = 
\begin{bmatrix}
2 & 0 & 0\\
0 & 2 & 0\\
1 & 0 & 2
\end{bmatrix}. 
$$

```{r}
B = matrix(c(
    2, 0, 0,
    0, 2, 0,
    1, 0, 2
  ), nrow = 3, byrow = TRUE)
B
```

Report on what happes if you try to run the following code to check if $B = V\Lambda V^{-1}$ is true? Why does it happen? 

```{r, eval = FALSE}
r <- eigen(B)
V <- r$vectors
Lambda <- diag(r$values)

all.equal(B + 0i, V %*% Lambda %*% solve(V))
```


